\documentclass[aspect=4:3]{nmd-slide}

\newcommand{\alg}[1]{\textsc{\textbf{#1}}}
\newcommand{\x}{\mathbf{x}}

\begin{document}

\begin{frame}

  \begin{center}
    {\Large Computational complexity}
    
    \medskip

    {\Large  of problems in }

    \medskip

    {\Large  3-dimensional topology}

    \vspace{2cm}

    {Nathan Dunfield}

    \smallskip

    University of Illinois

    \vspace{1cm}
    
    slides at: http://dunfield.info/preprints/
    
  \end{center}
\end{frame}

\begin{frame}

\textbf{[Novikov 1962]} \textsl{For $n \geq 5$, there does not exist an algorithm which solves:}

\vspace{0.2cm}

\alg{IsSphere:} Given a triangulated $M^n$ is it homeomorphic to $S^n$?

\vspace{1cm}

\textbf{Thm} (Geometrization + many results)  \\
\textsl{There is an algorithm to decide if two compact 3-mflds are homeomorphic.}

\vspace{3cm}

\textbf{Today:}  How hard are these 3-manifold questions?  How quickly can we solve them?

\vfill
\end{frame}

\begin{frame}
  \textbf{Decision Problems:}  Yes or no answer.  
  
  \bigskip
  
  \alg{Sorted:} Given a list of integers, is it sorted?

  \medskip

  \alg{SAT:} Given $p_1, \ldots p_n \in \F_2[x_1, \ldots, x_k]$ is there $\x \in \F_2^k$ with $p_i(\x) = 0$ for all $i$? 

  \medskip

  \alg{Unknotted:}  Given a planar diagram for $K$ in $S^3$ is $K$ the unknot?

  \medskip

  \alg{Invertible:} Given $A \in M_n(\Z)$ does it have an inverse in $M_n(\Z)$?

  \vspace{1.5cm}

  \textbf{P:} Decision problems which can be solved in polynomial time in the input size.

  \smallskip 

  \quad \alg{Sorted:} $O(\mbox{length of list})$

  \smallskip 

  \quad \alg{Invertible:} $O\big(n^{3.5} \log(\mbox{largest entry})^{1.1}\big)$
\end{frame}

\begin{frame}
  \textbf{NP:}  Yes answers have proofs that can be checked in
  polynomial time.  
  
  \medskip 
  \alg{SAT:} Given $\x \in \F_2^k$, can check all $p_i(\x) = 0$ in linear time. 

  \medskip

  \alg{Unknotted:} A diagram of the unknot with $c$ crossings can be unknotted
   in $O(c^{11})$ Reidemeister moves.  [Lackenby 2013]

\vspace{1.5cm}

\textbf{coNP:}  No answers can be checked in polynomial time.

 \medskip
 \alg{Unknotted:}  Yes, assuming the GRH [Kuperberg 2011]. 

\vspace{1.5cm}

\textbf{Conj:}  \alg{Unknotted} is in \textbf{P}.  

\end{frame}

\end{document}